---
title: "Notes on Learn Program WRDS with R"
date:  "`r Sys.Date()`"
author: Chris Zheng
email: chrizheng@vip.sina.com.cn
output:
  html_document:
     fig_caption: yes
     number_sections: yes
     toc: yes
     toc_depth: 4
     toc_float:
       collapsed: no
       smooth_scroll: no
     df_print: paged
     theme: cerulean
     highlight: pygments
---

```{r r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

# Learning from tourial of Wards
## Connect with WRDS

```{r connect_wrds}

library(RPostgres)

# Connect with WRDS database with my account
wrds <- dbConnect(Postgres(),
  host = "wrds-pgdata.wharton.upenn.edu",
  port = 9737,
  dbname = "wrds",
  sslmode = "require",
  user = "cz2003"
)
# withr::defer(
#   dbDisconnect(wrds)
# )
```

## Querying the Data set Structure (Metadata)
When working with WRDS data, it is often useful to examine the structure of
the dataset, before focusing on the data itself. WRDS data is organized by
vendor (such as crsp) and referred to as a library. Each library contains a
number of database tables or datasets (such as dsf), which contain the actual
data in tabular format with column headers called variables.

You can analyze the structure of the data through its metadata by querying the
information_schema table, as outlined in the following steps:

 1. Select a library to work with, and list all available datasets within that
   library

 2. Select a dataset, and list all available variables (column headers)
   within that dataset

**NOTE**: When referencing library and dataset names, you must
use all lowercase. This applies to both **information_schema.tables** and
**information_schema.columns** files.

```{r query_wrds_metadata}

# Determine the data libraries available at WRDS
res <- dbSendQuery(wrds, "select distinct table_schema
                   from information_schema.tables
                   where table_type ='VIEW'
                   or table_type ='FOREIGN TABLE'
                   order by table_schema")
wrds_libs <- dbFetch(res, n = -1)
dbClearResult(res)
head(wrds_libs)

# Determine the datasets within a given library:
res <- dbSendQuery(wrds, "select distinct table_name
                   from information_schema.columns
                   where table_schema='crsp'
                   order by table_name")
lib_tables <- dbFetch(res, n = -1)
dbClearResult(res)
head(lib_tables)

# Determine the variables (column headers) within a given dataset:
res <- dbSendQuery(wrds, "select column_name
                   from information_schema.columns
                   where table_schema='crsp'
                   and table_name='dsf'
                   order by column_name")
table_columns <- dbFetch(res, n = -1)
dbClearResult(res)
table_columns
```

## Query WRDS Data
Now that you've learned how to query the metadata and understand the structure
of the data, you are ready to query WRDS data directly. The dbSendQuery()
function can be used quite flexibly to pull WRDS data directly.

Unlike metadata queries, data queries reference both the data library and its
constituent dataset together in dot notation. For example, a data query for
the dataset **dsf** within the library **crsp** would use the syntax **crsp.dsf**.

**NOTE:** Unlike querying the PostgreSQL information_schema table, querying
database tables themselves (such as crsp.dsf) does not require that you adhere
to any specific case (i.e. upper- or lowercase). However, to keep things the
same across the board, WRDS recommends always using lowercase for referencing
libraries and their datasets.

```{r query_wrds_data}

# Query the crsp.dsf dataset

# Query all columns from a table
res <- dbSendQuery(wrds, "select * from crsp.dsf")
data <- dbFetch(res, n = 10)
dbClearResult(res)
head(data)

# Query specific variables from a table
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf")
data <- dbFetch(res, n = 10)
dbClearResult(res)
head(data)

# Query variables that meet certain criteria
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf
                   where askhi > 2500
                   and bidlo < 2000")
data <- dbFetch(res, n = 10)
dbClearResult(res)
head(data)

# Query by date
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf
                   where date = '2013-01-04'")
data <- dbFetch(res, n = -1)
dbClearResult(res)
head(data)

# Query by date range
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf
                   where date between '2013-01-07'
                   and '2013-01-08'")
data <- dbFetch(res, n = -1)
dbClearResult(res)
head(data)

# Query using multiple search criteria
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf
                   where date between '1960-01-01'
                   and '1980-01-01'
                   and askhi > 2500
                   and bidlo < 2000")
data <- dbFetch(res, n = -1)
dbClearResult(res)
head(data)

# Join and query two Compustat datasets
res <- dbSendQuery(wrds, "select a.gvkey, a.datadate, a.tic,
                   a.conm, a.at, a.lt, b.prccm, b.cshoq
                   from comp.funda a join comp.secm b
                   on a.gvkey = b.gvkey
                   and a.datadate = b.datadate
                   where a.tic = 'IBM'
                   and a.datafmt = 'STD'
                   and a.consol = 'C'
                   and a.indfmt = 'INDL'")
data <- dbFetch(res, n = -1)
dbClearResult(res)
head(data)
```

Graphing Data using R

```{r graph_wrds_data}

res <- dbSendQuery(wrds, "SELECT date,dji FROM djones.djdaily")
data <- dbFetch(res, n = -1)
plot(as.Date(data$date, "%Y-%m-%d"),
  data$dji,
  xlab = "date", ylab = "dji", type = "l", col = "red"
)
```

## Example Data Workflow using R

```{r wrds_example_workflow, eval= FALSE}

# Determine what data libraries are available at WRDS
res <- dbSendQuery(wrds, "select distinct table_schema
                   from information_schema.tables
                   where table_type ='VIEW'
                   or table_type ='FOREIGN TABLE'
                   order by table_schema")
data <- dbFetch(res, n=-1)
dbClearResult(res)
head(data)

# For the 'crsp' library, determine what datasets are available
res <- dbSendQuery(wrds, "select distinct table_name
                   from information_schema.columns
                   where table_schema='crsp'
                   order by table_name")
data <- dbFetch(res, n=-1)
dbClearResult(res)
head(data)

# For the Daily Stock File (dsf) dataset, determine what data variables (column
# headers) are available
res <- dbSendQuery(wrds, "select column_name
                   from information_schema.columns
                   where table_schema='crsp'
                   and table_name='dsf'
                   order by column_name")
data <- dbFetch(res, n=-1)
dbClearResult(res)
head(data)

# For the 'cusip', 'permno', 'date', 'bidlo', and 'askhi' data variables,. run a
# query limiting the result to the first 100 rows in the dataset
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf")
data <- dbFetch(res, n=100)
dbClearResult(res)
head(data)

# Run the query again, but filter the query for 'permno', limiting results to a
# single day
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf
                   where permno in (14593, 90319, 12490, 17778)
                   and date='2013-01-04'")
data <- dbFetch(res, n=100)
dbClearResult(res)
head(data)

# Determine high 'askhi' values by running a query to get a list of dates where
# 'permno' values posted an Ask Price over $2,000 between the years 2010 and
# 2013, no longer limiting the number of returned rows (as this is a pretty
# specific query)
res <- dbSendQuery(wrds, "select cusip,permno,date,bidlo,askhi
                   from crsp.dsf
                   where permno in (14593, 90319, 12490, 17778)
                   and date between '2010-01-01' and '2013-12-31'
                   and askhi > 2000")
data <- dbFetch(res, n=-1)
dbClearResult(res)
head(data)

# Only a single 'permno' posted a high Ask price during this date range. Open the
# search to all permnos that have ever posted an Ask Price over $2,000 in any
# date range (use distinct to return only one entry per matching permno). Since
# the query is against all permnos and the entire date range of the dataset, this
# query may take a little longer
res <- dbSendQuery(wrds, "select distinct permno
                   from crsp.dsf
                   where askhi > 2000")
data <- dbFetch(res, n=-1)
dbClearResult(res)
head(data)

# Retrieve all dates for which an Ask Price over $2000 was posted, along with
# the permnos that posted them. This will give a list of dates that match, with
# an additional entry for that date if additional permnos match as well. By
# 2011, there are two results
res <- dbSendQuery(wrds, "select distinct date,permno
                   from crsp.dsf
                   where askhi > 2000
                   order by date")
data <- dbFetch(res, n=-1)
dbClearResult(res)
head(data)

# Query for the highest Ask ever posted (searching only through Asks over
# $2000), on what date it posted, and which permno posted it. Use limit 1 to
# speed up the search since only the top value is desired:
res <- dbSendQuery(wrds, "select permno,askhi,date
                   from crsp.dsf
                   where askhi > 2000
                   order by askhi desc")
data <- dbFetch(res, n=1)
dbClearResult(res)
head(data)
```

